### 2025-08-05 conteudo/tool/architecture.tex
Review tool

Regarding the programming language's function extractor, we approached this problem so that anyone 
comfortable with a language can adapt our approach to their specific programming language. 
Usually, code blocks can be easily extracted from source code files, independently of whether the 
code blocks are defined by curly brackets or by indentation, and it is possible to infer if 
they represent a function's body given their depth in the code context and a few lines before the 
start of the code block. We followed this approach to implement the function extractor for the 
supported languages. As an alternative approach, we can build the syntax tree \citep{compiler} 
of the programming language and use it to extract the functions, which is a more complex task than 
our approach. For the programming languages that we support, we implemented them as follows:

@@

For a source file in the C programming language, we extract the code blocks with depth 0. 
As per the programming language’s grammar, functions are defined in a global context at depth 0. 
To extract the code blocks, we first need to identify the portions of code that are comments or 
\textit{\#define} directives and mark them as irrelevant. After that, it is simple work to find 
the curly brackets in depth 0; with that, we find the body of the functions.

@@

For the Java programming language, we have decided to implement a more straightforward approach 
that does not support the programming language completely, as the only Java codebase explored in 
this work is the BigCloneBench \citep{bigclonebench}, and the approach implemented covers the 
needs of this codebase.

@@

As per Java’s grammar, functions are defined inside classes and can exist at different depths. 
As Java accepts inner classes, this approach does not work with inner classes. The challenge 
to implementing a more generic approach is that other elements of depth two or deeper, such 
as for-loops, can follow the closing parenthesis pattern, necessitating a better analysis of 
these elements. The BigCloneBench \citep{bigclonebench} does not contain inner classes.

@@

The temporary codebase is a transformation of the input codebase performed by the Function Parser. Each function in the input codebase is represented in the temporary codebase as three files. Descriptions of these files are provided below, with examples shown in Figure \ref{fig:transform}.

\begin{itemize}
	\item \textbf{Header file}: This file contains the signature of the function it represents.
	\item \textbf{Source file}: This file contains the body of the function it represents.
	\item \textbf{Info file}: This file contains metadata about the function, including the function name, the file that contains the function, the file's relative path to the input codebase, the line where the function's signature starts, the line where the function body begins, the line where the function's body ends, and whether there is a line break between the end of the function's signature and the start of the function's body.
\end{itemize}

\subsection{Code Duplication Finder}

\label{subsec:code_finder}

@@

For the text similarity method, we treat the source code files 
as text and apply the TF-IDF vector embedding method implemented by the 
Gensim library \citep{gensim}, then compute cosine similarity as the 
similarity metric.

@@

\cite{platistool} implemented this code duplication detection method and 
released it under the MIT license \citep{mitlicense}. We adapted his 
implementation to fit our tool's expected input and output formats. 
Switching the duplication detection method is feasible, similar to how 
we adapted Plates' implementation.

@@

The Preprocessor is a procedure executed by the user once per codebase. 
The Preprocessor takes the input codebase and a minimum similarity metric value for function pairs considered duplicates. 
It then runs the Function Parser and the Code Duplication Finder to create a temporary codebase and the Code Duplication Database. 
Setting a minimum similarity metric reduces the Code Duplication Database size, optimizing memory usage and the computational 
cost of the Query Responder step. For example, if this parameter does not exist and we work with a codebase of $10000$ functions, 
each with a 50-character relative path and function name, the Code Duplication Database would reach approximately $5000^2 \times 50 ~= 5$ gigabytes. 
As most function pairs are unlikely to be duplicated in large codebases, this file size can be significantly reduced.

@@

The Query Responder step is the component that the user executes multiple times to consult information 
about duplicated functions in a codebase processed by the Preprocessor. 
It utilizes the Temporary Codebase and the Code Duplication Database as inputs to answer user requests, 
such as querying for duplicated functions above a certain similarity threshold or providing counts of such functions.
This component was developed to be extensible, allowing other query functionalities to be easily added. 
The specific functionalities provided to the user by the Query Responder, which include the 
Duplication Explorer, Function Information, and Duplication Report, are described in Section \ref{subsec:func}.

>>>>

Regarding the programming language function extractor, we approached this problem so that anyone 
comfortable with a language can adapt our approach to their specific programming language. 
Usually, code blocks can be easily extracted from source code files, independently of whether the 
code blocks are defined by curly brackets or by indentation, and it is possible to infer if 
they represent a function body given their depth in the code context and a few lines before the 
start of the code block. We followed this approach to implement the function extractor for the 
supported languages. As an alternative approach, we can build the syntax tree~\citep{compiler} 
of the programming language and use it to extract the functions, which is a more complex task than 
our approach. For the programming languages that we support, we implemented them as follows:

@@

For a source file in the C programming language, we extract the code blocks with depth 0. 
As per the programming language grammar, functions are defined in a global context at depth 0. 
To extract the code blocks, we first need to identify the portions of code that are comments or 
\textit{\#define} directives and mark them as irrelevant. After that, it is simple work to find 
the curly brackets in depth 0; with that, we find the body of the functions.

@@

For the Java programming language, we have decided to implement a more straightforward approach 
that does not support the programming language completely, as the only Java codebase explored in 
this work is the BigCloneBench~\citep{bigclonebench}, and the approach implemented covers the 
needs of this codebase.

@@

As per Java grammar, functions are defined inside classes and can exist at different depths. 
As Java accepts inner classes, this approach does not work with inner classes. The challenge 
to implementing a more generic approach is that other elements of depth two or deeper, such 
as for-loops, can follow the closing parenthesis pattern, necessitating a better analysis of 
these elements. The BigCloneBench~\citep{bigclonebench} does not contain inner classes.

@@

The temporary codebase is a transformation of the input codebase performed by the Function Parser. Each function in the input codebase is represented in the temporary codebase as three files. Descriptions of these files are provided below, with examples shown in Figure~\ref{fig:transform}.

\begin{itemize}
	\item \textbf{Header file}: This file contains the signature of the function it represents.
	\item \textbf{Source file}: This file contains the body of the function it represents.
	\item \textbf{Info file}: This file contains metadata about the function, including the function name, the file that contains the function, the file relative path to the input codebase, the line where the function signature starts, the line where the function body begins, the line where the function body ends, and whether there is a line break between the end of the function signature and the start of the function body.
\end{itemize}

\subsection{Code Duplication Finder}
\label{subsec:code_finder}

@@

For the text similarity method, we treat the source code files 
as text and apply the TF-IDF vector embedding method implemented by the 
Gensim library~\citep{gensim}, then compute cosine similarity as the 
similarity metric.

@@

\cite{platistool} implemented this code duplication detection method and 
released it under the MIT license. We adapted his 
implementation to fit our tool expected input and output formats. 
Switching the duplication detection method is feasible, similar to how 
we adapted Plates' implementation.

@@

The Preprocessor is a procedure executed by the user once per codebase. 
The Preprocessor takes the input codebase and a minimum similarity metric value for function pairs considered duplicates. 
It then runs the Function Parser and the Code Duplication Finder to create a temporary codebase and the Code Duplication Database. 
Setting a minimum similarity metric reduces the Code Duplication Database size, optimizing memory usage and the computational 
cost of the Query Responder step. For example, if this parameter does not exist and we work with a codebase of $10,000$ functions, 
each with a 50-character relative path and function name, the Code Duplication Database would reach approximately $5,000^2 \times 50 ~= 5$ gigabytes. 
As most function pairs are unlikely to be duplicated in large codebases, this file size can be significantly reduced.

@@

The Query Responder step is the component that the user executes multiple times to consult information 
about duplicated functions in a codebase processed by the Preprocessor. 
It utilizes the Temporary Codebase and the Code Duplication Database as inputs to answer user requests, 
such as querying for duplicated functions above a certain similarity threshold or providing counts of such functions.
This component was developed to be extensible, allowing other query functionalities to be easily added. 
The specific functionalities provided to the user by the Query Responder, which include the 
Duplication Explorer, Function Information, and Duplication Report, are described in Section~\ref{subsec:func}.
